\dppTitle

### 2014-06-02 manuscripts/dpp-msbayes/reviews/response.tex

    \begin{quote}
        Each \divTime{} within \divTimeVector is \iid as $\divTime{} \sim
        Gamma(\gammaShape{\divTime{}}, \gammaScale{\divTime{}})$, \change{
        where \gammaShape{\divTime{}} and \gammaScale{\divTime{}} are specified
        by the investigator}.
    \end{quote}
}

\revcomment{
    The ABC part is rather standard, except for the rather strange feature that
    the divergence times are used to construct summary statistics (p.10). I
    find it strange because these times are not observed for the actual data.
    So I must be missing something, which should be spelled out more clearly.
    (And I object to the sentence ``I have introduced a new approximate-Bayesian
    model choice method.'' (p.21) since there is no new ABC technique, simply a
    different form of prior.)
}
\response{
    Reviewer 2 is correct that there is no new ABC numerical machinery
    presented in the manuscript, only a new model. To make this clearer,
    I have changed the sentence to:
    \begin{quote}
        I introduced a new model for estimating shared divergence histories
        across taxa from DNA sequence data within an approximate-Bayesian
        model-choice framework.
    \end{quote}

    Also, the intuition of Reviewer 2 is correct, I do not use the divergence
    times to calculate summary statistics. I mentioned the divergence times in
    the description of the ABC algorithm with the hope of making it clear that
    the times are scaled (see Equation (12)) prior to the simulation of the
    data (from which the summary statistics are calculated). This scaling is
    simply to go from units proportional to time, to units that are
    proportional to the expected number of mutations. Because, this is already
    described earlier in the description of the model, I have removed the
    re-description of this scaling from the description of the ABC algorithm.

@@

        The magnitude of the affect of this violation of exchangeability is not
        known.  Huang et al.\ \cite{Huang2011} demonstrated that the reordering
        of the summary statistic vectors can greatly increase the method's
        tendency to infer a single divergence event.  By definition, if the
        summary statistic vectors were exchangeable, the reordering would not
        change the likelihood or posterior (barring sampling error).  Thus, the
        results of Huang et al.\ \cite{Huang2011} suggest the reordering of the
        statistics is potentially introducing sizeable error to the analysis.}
    \end{quote}
}

\revcomment{
    Quoting [43] [[citation 43 from the manuscript is now 10 in the current
    document]] as concluding that ABC model choice is biased is somewhat
    misleading: the warning therein was that Bayes factors and posterior
    probabilities resulting from those ABC steps could be quite unrelated with
    those based on the whole dataset. That the same authors concluded that the
    proper choice of summary statistics leads to a consistent model choice
    shows ABC model choice is not necessarily ``biased''.
}
\response{
    This is a very good point, thank you. I have expanded this section to be
    clearer about the findings of \cite{Robert2011} and how they relate to the
    methods discussed in the paper:
    \begin{quote}
        I also urge caution when using \dppmsbayes due to the lack of
        theoretical validation of Bayesian model choice when the full data are
        replaced by summary statistics that are insufficient for discriminating
        across models under comparison \cite{Robert2011}, which is certainly
        the case here.  Robert et al.\ \cite{Robert2011} demonstrated that ABC
        estimates of model posterior probabilities can be inaccurate when such
        across-model insufficient statistics are used.
    \end{quote}
}

### 2014-06-02 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding Dryad citation to dpp ms.

\subsection{An empirical application}
I also assessed the behavior of the newly implemented models when applied to
the empirical dataset of Oaks et al.\ \cite{Oaks2012}, which is comprised of
sequence data from 22 pairs of taxa from the Philippine Islands (Drayad DOI:
10.5061/dryad.5s07m).
I analyzed these data under five different models, which are detailed in
Table~\ref{tabEmpiricalModels}.
All of these models except one (\empModelDPPSimple) have six free demographic
parameters per pair of taxa (\ancestralTheta{},
\descendantTheta{1}{},
\descendantTheta{2}{},
\bottleTime{},
\bottleScalar{1}{}, and
\bottleScalar{2}{}),
in addition to the $\popSampleSize{i}{} - 1$ coalescent times.
Three of these models use a Dirichlet-process prior on divergence models:
\empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
The \empModelDPP model represents the priors that Oaks et al.\ \cite{Oaks2012}
would have selected to reflect their prior uncertainty about the parameters of
the model if provided the more flexible distributions that are now implemented.
To assess prior sensitivity, the \empModelDPPInform model uses a more
informative exponentially distributed prior on divergence times, but otherwise
is identical to \empModelDPP.
To assess sensitivity to parameterization, I also applied the simplest
possible model under the new implementation (\empModelDPPSimple) with only a
single demographic parameter (\myTheta{}) per taxon pair, in addition to the
$\popSampleSize{i}{} - 1$ coalescent times.
I also applied the original \msb model (\empModelOld) with priors selected to
make the results directly comparable to those of the \empModelDPP model;
the uniform prior on divergence times was selected to have the same variance as
the exponential prior of the \empModelDPP model, and the prior on population
size was selected to have the same mean so that the models are on the same
timescale.
I also applied a model with a uniform distribution over divergence models
(\empModelUniform).
For each of these models, I simulated 2\e{7} samples from the prior, and
retained an approximate posterior of the 10,000 samples with the smallest
Euclidean distance from the summary statistics calculated from the empirical
sequence alignments.

>>>>

\subsection{An empirical application}
I also assessed the behavior of the newly implemented models when applied to
the empirical dataset of Oaks et al.\ \cite{Oaks2012}, which is comprised of
sequence data from 22 pairs of taxa from the Philippine Islands
(\cite{Oaks2012dryad}; Dryad DOI: 10.5061/dryad.5s07m).
I analyzed these data under five different models, which are detailed in
Table~\ref{tabEmpiricalModels}.
All of these models except one (\empModelDPPSimple) have six free demographic
parameters per pair of taxa (\ancestralTheta{},
\descendantTheta{1}{},
\descendantTheta{2}{},
\bottleTime{},
\bottleScalar{1}{}, and
\bottleScalar{2}{}),
in addition to the $\popSampleSize{i}{} - 1$ coalescent times.
Three of these models use a Dirichlet-process prior on divergence models:
\empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
The \empModelDPP model represents the priors that Oaks et al.\ \cite{Oaks2012}
would have selected to reflect their prior uncertainty about the parameters of
the model if provided the more flexible distributions that are now implemented.
To assess prior sensitivity, the \empModelDPPInform model uses a more
informative exponentially distributed prior on divergence times, but otherwise
is identical to \empModelDPP.
To assess sensitivity to parameterization, I also applied the simplest
possible model under the new implementation (\empModelDPPSimple) with only a
single demographic parameter (\myTheta{}) per taxon pair, in addition to the
$\popSampleSize{i}{} - 1$ coalescent times.
I also applied the original \msb model (\empModelOld) with priors selected to
make the results directly comparable to those of the \empModelDPP model;
the uniform prior on divergence times was selected to have the same variance as
the exponential prior of the \empModelDPP model, and the prior on population
size was selected to have the same mean so that the models are on the same
timescale.
I also applied a model with a uniform distribution over divergence models
(\empModelUniform).
For each of these models, I simulated 2\e{7} samples from the prior, and
retained an approximate posterior of the 10,000 samples with the smallest
Euclidean distance from the summary statistics calculated from the empirical
sequence alignments.
